Cosmological constraints from the redshift dependence of the galaxy angular 2-point correlation function

### 2017-04-01 draft/draft1.tex
updates

\author{Xiao-Dong Li, }
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\author{Cheng Cheng, }
\affil{Kavli Institute for Theoretical Physics China, Institute of Theoretical Physics, Chinese Academy of Sciences, Zhong Guan Cun Street 55\#, Beijing, 100190, P.R. China}
\affil{University of Chinese Academy of Sciences, P.R. China}
\author{Changbom Park,}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\author{Hyunbae Park, Cristiano G. Sabiu, }
\affil{Korea Astronomy and Space Science Institute, Daejeon 305-348, Korea}
%\author{David H. Weinberg,}
%\affil{Department of Astronomy and CCAPP, The Ohio State University, 140 West 18th Avenue, Columbus, OH 43210, USA}
%\author{Donald P. Schneider,}
%\affil{Department of Astronomy and Astrophysics, The Pennsylvania State University, 
%University Park, PA 16802 }
%\affil{Institute for Gravitation and the Cosmos, The Pennsylvania State University, 
%University Park, PA 16802 }
\author{Juhan Kim\altaffilmark{1},}
\affil{Center for Advanced Computation, Korea Institute for Advanced Study, 85 Hoegi-ro, Dongdaemun-gu, Seoul 130-722, Korea}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\and
\author{Sungwook E. Hong}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}

%\altaffiltext{1}{xiaodongli@kias.re.kr}
%\altaffiltext{2}{cbp@kias.re.kr}
%\altaffiltext{3}{csabiu@kasi.re.kr}
%\altaffiltext{4}{\bf Hyunbae: your email}
\altaffiltext{1}{Corresponding Author: kjhan@kias.re.kr}

@@

\begin{abstract}
We propose to use the redshift dependence of the galaxy angular 2-point correlation function (2pCF)
for a purpose of cosmological constraints.
When an incorrect cosmological model is assumed for the coordinate transformation from redshift space to comoving space, 
it produces {\it redshift-dependent mis-scaling} in the constructed galaxy distribution.
As a result, we observe a redshift-dependent rescaling of the measured galaxy angular 2pCF.
We test our method on the HR4 mock galaxies having hundreds of millions of galaxies at redshifts 
$0\leq z\leq2$.
We find the redshift evolution of shape of angular 2pCF, as a function of scale,
is sensitive to the adopted cosmological parameters,
and is relatively insensitive to the gravitational growth of structure,
the galaxy bias, and the redshift-space distortion.
Analyzing the 2pCF shape on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$, 
we derive tight constraints on $\Omega_m$ and $w$ with 68.3\% CL uncertainties of 0.015 and 0.05.
%Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future large scale structure surveys, especially the photometric surveys such as DES, LSST, 
to derive tight cosmological constraints.
This work is a continuation of our previous works \citep{Li2014,Li2015,Li2016}
as a strategy to constraining cosmological parameters based on some nearly redshift-invariant physical quantity;
they are all powerful in constraining cosmology and have a lot of merit,
e.g., are less affected by the systematic effects introduced by non-linear clustering or RSD,
do not require accurate analytical modeling of the galaxy correlation function,
and are applicable on relatively small clustering scales.
\end{abstract}

@@

In \cite{Li2014,Li2015,Li2016} we developed a novel strategy to probe $D_A$ and $H$ from LSS data.
%proposed a series of methods to probe the geometric information encoded in the LSS data.
We found that the AP or volume effect in general introduces a geometric distortion that evolves significantly with redshift;
this phenomenon can be utilized to distinguish the AP effect from the other effects, such as the redshift-space distortion (RSD).
The radial distances of galaxies are inferred from the measured redshifts, which are distorted by the peculiar motion of galaxies, 
leading to apparent distortion in the redshift space galaxy distribution \citep{FOG,Kaiser1987,Ballinger1996}.
This is the major systematic effect limiting our ability to probe the geometry of LSS from galaxy distribution.
We found its influence is significantly reduced when studying the redshift dependence of the distortion.

\cite{Li2014} attempted to use the {\it galaxy density gradient field} 
to characterize the redshift evolution of the LSS distortion. 
We tested the idea on the Horizon Run 3 (HR3) N-body simulation \citep{horizonrun},
showing that the redshift evolution of the gradient field 
can be used to infer unbiased constraints on cosmological parameters and that it is, largely, insensitive to RSD.
The same topic was revisited in \cite{Li2015}, 
using a slightly modified version of the usual anisotropic galaxy two-point correlation function (2pCF), $\xi(\mu)$, 
where $\mu\equiv \cos \theta$ and $\theta$ is the direction between the connection of the galaxy pairs and the line of sight.
When an incorrect cosmological model is adopted, 
the shape of $\xi(\mu)$ is altered due to the anisotropic distortion of the LSS (the AP effect), 
and the amplitude is shifted due to the change in the comoving volume elements;
both effects have significant redshift dependence.
Although the RSD effect significantly distorts $\xi(\mu)$, 
it exhibits much less redshift evolution.
The method was applied to BOSS DR12 galaxies \citep{Li2016} to probe the redshift dependence of the AP effect, 
and we arrive at tight cosmological constraints comparable to mainstream probes of SNIa, BAO, CMB.

@@

\scr{When the correct cosmology is adopted the cosmic scale is correctly estimated,
the main factor leads to redshift evolution of galaxy distribution is the gravitational growth of structure.
With decreasing redshift, we clearly see the clusters and filaments form and grow.
On the other hand, when the wrong cosmology is adopted, there is an artificial scaling of distances in the constructed map.
The separations among galaxies are over-estimated by 
$25.6\%,47.3\%,62.2\%,71.7\%$ at redshifts of $0.5,1.0,1.5,2.5$,
leading to a clear redshift evolution of sizes of structures.}

@@

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig3.eps}
    \includegraphics[width=18cm]{fig3_b.eps}
   }
   \caption{\label{fig_diffz}
  The angular 2pCF ($\omega_{r_\perp}$, left panels), the shape of the angular 2pCF ($\hat{\omega}_{r_\perp}$, middle panels),
  and the evolution with respect to the $z=0$ value ($\hat{\omega}_{r_{\perp}}(z=z_i) - \hat{\omega}_{r_{\perp}}(z=0)$, right panels). 
  Upper panels show that in the correct cosmology the redshift evolution of the shape of 2pCF is rather small, 
  whereas the lower panels show that, in a wrong cosmology $\Omega_m = 0.05,w=-1.5$, the redshift evolution of $\hat{\omega}_{r_\perp}$ is very significant.
   }
\end{figure*}

The upper-left panel of Figure \ref{fig_diffz} displays the angular 2pCF measured from HR4 mock galaxies. % (multiplied by scale $r$) in the five snapshot.
We multiply $\omega$ by the separation $r_\perp$ to have similar size of statistical uncertainty on all scales.

Among different redshifts there is large variation in the amplitude of $\omega(r_\perp)$.
The amplitude is proportional to the clustering strength, which is affected by the gravitational growth of structures and the galaxy bias.
The amplitude is highest at $z=0$ where the structures experienced the most growth.
$\omega(r_\perp)$ also increases with redshift at $z>1$ with increasing redshift 
as a reason of increasing galaxy bias.
%leading to stronger clustering strength and higher amplitude of 2pCF.
%On the contrary, at higher redshift we find $\omega(z=2)>\omega(z=1.5)>\omega(z=1)$,
%The trend is different from $z=1,1.5,2$,
%simply because by applying a uniform mass cut we are selecting more biased galaxies at higher redshift.

Although there is large variation in amplitude between the different redshifts bins, 
%The shape of $r_\perp\omega$ also evolves with redshift.
the shape of $r_\perp\omega$  remains similar at all redshifts;
in general it peaks at $r\sim9 h^{-1}\rm Mpc$ and monotonically drops or increases at larger or smaller scales.
The only exception is the small enhancement at $r\lesssim 2 h^{-1}\rm Mpc$,
which is caused by the non-linear growth of structures 
and is much more significant at lower redshift.

@@

In the upper-middle panel, the overlapping of $\hat\omega$
clearly show the minimal redshift evolution of the shape.
In the upper-right panel, 
we further show the residual evolution at high redshifts 
with respect to $z=0$.
There is only 1-4\% enhancement 
at $r<10h^{-1}\rm Mpc$,
and $<1.5\%$ suppression at $r>25 h^{-1}\rm Mpc$.
The trend is monotonic with redshift.

@@

\subsection{Galaxy angular 2pCF: Cosmological Effect }

The angular 2pCF of galaxy positions constructed in a wrongly assumed cosmology $\Omega_m=0.05,\ w=-1.5$
is displayed in the lower panels of Figure \ref{fig_diffz}.
The mis-scaling uniformly shifts the clustering pattern on all scales, 
leading to a biased $\hat \omega_{r_\perp}$, which is 
related with the true $\hat \omega$ as
\begin{equation}
 \hat \omega_{r_\perp,\rm wrong}(r) = \hat \omega_{r_\perp, \rm correct}(\alpha_{\perp} r),
\end{equation}
a simple consequence of the fact that the clustering pattern at scale $r$ is rescaled to $\alpha_{\perp} r$.
The redshift evolution of $\alpha_{\perp}$ leads to redshift evolution of 
$\hat \omega_{r_\perp, \rm correct}$,
%Lower panels of Figure \ref{fig_diffz} shows the $r_\perp\omega$ in five 
%snapshots, in case that the cosmology $\Omega_m=0.05,\ w=-1.5$ 
%is adopted to construct the galaxy distribution (the right panels of Figure \ref{fig_scatter}).
which is displayed in the lower-middle panel.
The upscaling of comoving distance leads to a stretched shape.
When increasing redshift,
the peak location is shifted from $\sim9h^{-1} \rm Mpc$ to $\sim15 h^{-1}\rm Mpc$ at $z=2$,
as a result of the fact that the angular separation is upscaled by $71.7\%$.
Correspondingly, the right panel shows that, 
compared with $\hat \omega_{r_\perp}$ at $z=0$ there is a 20-40\% change at higher redshifts.

@@

We vary the mass cut and check the effect of galaxy bias on $\hat \omega_{r_\perp}$.
The upper panels of Figure \ref{fig_sys} displays the 2pCF measured from galaxies within a $3175\times3175\times105$ $(h^{-1}\rm Mpc)^3$ volume, taken from the $z=0$ snapshot.
Four values of minimal mass limits, $3\times 10^{11} M_{\odot}$, $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$,
are imposed to create three \scr{four?} samples with different galaxy bias.
The measured angular 2pCF are displayed.

Different mass cuts result in large variation in the amplitude of the 2pCF (left panel),
while the shape of the \scr{normalised} 2pCF remains less affected (middle panel).
Compared with the sample of $M>3\times 10^{11} M_{\odot}$,
samples with mass limits of $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$
have amplitude of 2pCFs enhanced by $10\%,\ 50\%$ and $100\%$, respectively, 
while the change in the shape is only $0.5\%,\ 2\%$ and $4\%$.
The effect of galaxy bias becomes much smaller by utilizing the shape of the 2pCF rather than the amplitude.

@@

The effect of RSD is much milder in this work. 
The angular positions of galaxies are not shifted by RSD at all;
the only effect of RSD enter in the procedure of 
splitting galaxies into redshift shells.
The galaxies observed in a survey are split into shells of subsamples, 
with different redshift ranges, for the 2pCF measurement.
RSD distorts the galaxy redshift and as a result some galaxies 
(especially those close to the boundaries of shells) are assigned to the wrong redshift shells.

We shift the radial coordinates of galaxies according to the relation 
\begin{equation}\label{eq:zvpeu}
\Delta z = (1+z) \frac{v_{{\rm Z}}}{c},
\end{equation}
where $v_{{\rm Z}}$ (Z is treated as the radial direction in this analysis) 
is the galaxy peculiar velocity in the LOS direction.
This lead to misclassification of some galaxies when we split the box into slices.
%Then we split the whole box into 30 slices with thickness $105h^{-1}\rm Mpc$, 
%and measure the angular 2pCF.

The lower panel of Figure \ref{fig_sys} shows the RSD effect on the 2pCF.
Comparing this plot with Figure \ref{fig_diffz} one can clearly see the effect of RSD.
The amplitude of the measured 2pCF is enhanced by $\sim 10\%$ when considering the RSD effect 
and the slop of $\hat\omega_{r_\perp}$ is suppressed.
The redshift evolution of $\hat\omega_{r_\perp}$ is still small, remain $\lesssim4\%$ at $0.5\leq z \leq 2$.
So RSD should not very affect the application of our method as long as it is precisely modeled and corrected.
%If looking at the redshift evolution,
%the curves $\hat{r_\perp\omega} - \hat{r_\perp\omega}(z=0)$ displayed in the right panel,
%they changed a lot compared with the case of no RSD effect (the right panel of Figure \ref{fig_diffz}),
%but do not become much larger. %But RSD effect does not introduce large redshfit evolution, although the shape of  is

Similar to RSD, redshift errors of galaxies
also results in fuzzy boundaries of shells and misclassification of galaxies.
This effect should also be properly quantified in real observations.

@@

Adopting the correct cosmological model will resulting in minimal evolution of the normalised 2pCF.
Thus we proceed to construct a quantitative likelihood estimator.
The procedure here is very similar to \cite{Li2014,Li2015,Li2016}.
We compare the high redshift $\hat\omega_{r_\perp}$ and the lowest redshift measurement,
\begin{equation}
 \delta \hat{r_\perp\omega}(z_i,z_1) \equiv \hat{r_\perp\omega}(z_i) - \hat{r_\perp\omega}(z_1),
\end{equation}
and design the likelihood function to quantify
the redshift evolution of $\hat\omega_{r_\perp}$:
\begin{equation}\label{eq:chisq1}
\chi^2\equiv \sum_{i=2}^{n_{z}} \sum_{j_1=1}^{n_{r}} \sum_{j_2=1}^{n_{r}} {\bf p}(z_i,r_{j_1}) ({\bf Cov}_{i}^{-1})_{j_1,j_2}  {\bf p}(z_i,r_{j_2}),
\end{equation}
where $n_z$ is the number of redshifts which is 5 in this analysis, 
$n_r$ is number of binning in $\hat{\omega}_{r_\perp}(r_\perp)$,
which is 35 since we have $r_\perp$ bins 
with width $1h^{-1}\rm Mpc$ in a range of $5h^{-1}{\rm Mpc}\leq r\leq 40h^{-1}{\rm Mpc}$.
${\bf p}(z_i,r_{j})$ is the redshift evolution of the angular correlation function shape with systematic effects subtracted
\begin{eqnarray}\label{eq:bfp}
 {\bf p}(z_i,r_{j}) \equiv&\ \delta \hat{\omega}(z_i,z_1,r_j) - \delta \hat{\omega}_{r_\perp,\rm sys}(z_i,z_1,r_j)
\end{eqnarray}
${\bf Cov}_i$ is the covariance matrix estimated from the ${\bf p}(z_i,r_{j})$ measured from 120 subsamples.
%It is difficult to conduct the 2pCF of the whole snapshot box with hundreds of millions of galaxies.
%We split the box into 120 subsamples with $X,Y,Z$ sizes of $1575h^{-1}{\rm Mpc},1575h^{-1}{\rm Mpc}, 105h^{-1}{\rm Mpc}$
%and measure the angular 2pCF in the $X-Y$ plan in these subsamples. 
%The average of the 2pCFs are taken as the 2pCF of the whole box.
%The covariance matrix is also estimated from the redshift evolution of these subsamples.
As mentioned previously, for a robust estimation of the covariance matrix, 
we always compare slices at different locations to 
include the cosmic variance.

@@

Figure \ref{fig_cosmo} displays the 2pCF and the likelihood values when adopting 
the four incorrect cosmologies used in Figure \ref{fig_xyquan}.
For the cosmologies $\Omega_m=0.4,w=-1$ and $\Omega_m=0.26,w=-0.5$,
the compression shifts the large scale clustering patterns into small scales
at higher redshift and make the 2pCFs steeper.
%and also becomes more significant at higher redshift.
%We get 2pCFs with steeper slope and the higher redshift the steeper.
For cosmologies $\Omega_m=0.15,w=-1$ and $\Omega_m=0.26,w=-1.5$,
the stretch of structure leads to shallower slope of $\hat{\omega}_{r_\perp}$ at higher redshift.
%the shallower.
In all cosmologies there is significant detection of redshift evolution.
We compute the $\chi^2$ values according to Equation \ref{eq:chisq1}, 
and find that these cosmologies are disfavored 
at $\gtrsim30\sigma$ CL.

>>>>

\author{Xiao-Dong Li, Changbom Park, }
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\author{Cristiano G. Sabiu, Hyunbae Park, }
\affil{Korea Astronomy and Space Science Institute, Daejeon 305-348, Korea}
%\author{David H. Weinberg,}
%\affil{Department of Astronomy and CCAPP, The Ohio State University, 140 West 18th Avenue, Columbus, OH 43210, USA}
%\author{Donald P. Schneider,}
%\affil{Department of Astronomy and Astrophysics, The Pennsylvania State University, 
%University Park, PA 16802 }
%\affil{Institute for Gravitation and the Cosmos, The Pennsylvania State University, 
%University Park, PA 16802 }
\author{Cheng Cheng, }
\affil{Kavli Institute for Theoretical Physics China, Institute of Theoretical Physics, Chinese Academy of Sciences, Zhong Guan Cun Street 55\#, Beijing, 100190, P.R. China}
\affil{University of Chinese Academy of Sciences, P.R. China}
\author{Juhan Kim\altaffilmark{1},}
\affil{Center for Advanced Computation, Korea Institute for Advanced Study, 85 Hoegi-ro, Dongdaemun-gu, Seoul 130-722, Korea}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}
\and
\author{Sungwook E. Hong}
\affil{School of Physics, Korea Institute for Advanced Study, 85 Heogiro, Dongdaemun-gu, Seoul 130-722, Korea}

%\altaffiltext{1}{xiaodongli@kias.re.kr}
%\altaffiltext{2}{cbp@kias.re.kr}
%\altaffiltext{3}{csabiu@kasi.re.kr}
%\altaffiltext{4}{\bf Hyunbae: your email}
\altaffiltext{1}{Corresponding Author: Cris, Add your email!!!}

@@

\begin{abstract}
We propose to use the redshift dependence of the galaxy angular 2-point correlation function (2pCF)
to constrain cosmological parameters.
When an incorrect set of parameters is assumed for the coordinate transformation from redshift space to comoving space, 
it produces {\it redshift-dependent mis-scaling} in the constructed galaxy distribution.
As a result, we observe a redshift-dependent rescaling of the measured galaxy angular 2pCF.
We test our method on the Horizon Run 4 N-body simulation snapshot data, 
which contains hundreds of millions of galaxies at redshifts 
0, 0.5, 1, 1.5 and 2
We find the shape of angular 2pCF, as a function of scale,
exhibits evident redshift evolution when an incorrect set of cosmological parameters adopted,
and is relatively insensitive to the other effects such as gravitational growth of structure,
the galaxy bias, and the redshift-space distortion.
Analyzing the 2pCF shape on scales $5  h^{-1} {\rm Mpc} \leq r_\perp \leq 40 h^{-1} {\rm Mpc}$, 
we derive tight constraints on $\Omega_m$ and $w$ with 68.3\% CL uncertainties of 0.015 and 0.05.
%Our method is complementary to current works using 2pCF to constrain cosmological parameters.
The method could be applicable to future large scale structure surveys, especially the photometric surveys such as DES, LSST, 
to derive tight cosmological constraints.
This work is a continuation of our previous works \cite{Li2014,Li2015,Li2016},
which use a strategy to constraining cosmological parameters based on some redshift-invariant physical quantity;
they have a lot of merit,
e.g., lead to powerful constraints, are less affected by the systematic effects introduced by non-linear clustering or RSD,
do not require accurate analytical modeling of the galaxy correlation function,
are applicable on relatively small clustering scales,
and are applicable to many future LSS surveys.
\end{abstract}

@@

In \cite{Li2014,Li2015,Li2016} we developed a novel strategy to probe $D_A$ and $H$ from the LSS data.
%proposed a series of methods to probe the geometric information encoded in the LSS data.
We found that the AP or volume effect in general introduces a geometric distortion that evolves significantly with redshift;
this phenomenon can be utilized to distinguish the AP effect from the other effects, such as the redshift-space distortion (RSD).
The radial distances of galaxies are inferred from the measured redshifts, which are distorted by the peculiar motion of galaxies, 
leading to apparent distortion in the redshift space galaxy distribution \citep{FOG,Kaiser1987,Ballinger1996}.
This is the major systematic effect limiting our ability to probe the geometry of LSS from galaxy distribution.
We found its influence is significantly reduced when studying the redshift dependence of the distortion.

\cite{Li2014} attempted to use the {\it galaxy density gradient field} 
to characterize the redshift evolution of the LSS distortion. 
We tested the idea on the Horizon Run 3 (HR3) N-body simulation \citep{horizonrun},
showing that the redshift evolution of the gradient field 
can be used to infer unbiased constraints on cosmological parameters and that it is, largely, insensitive to RSD.
The same topic was revisited in \cite{Li2015}, 
using a slightly modified version of the usual anisotropic galaxy two-point correlation function (2pCF), $\xi(\mu)$, 
where $\mu\equiv \cos \theta$ and $\theta$ is the direction between the connection of the galaxy pairs and the line of sight.
When an incorrect cosmological model is adopted, 
the shape of $\xi(\mu)$ is altered due to the anisotropic distortion of the LSS (the AP effect), 
and the amplitude is shifted due to the change in the comoving volume elements;
both effects have significant redshift dependence.
Although the RSD effect significantly distorts $\xi(\mu)$, 
it exhibits much less redshift evolution.
The method was applied to BOSS DR12 galaxies \citep{Li2016} to probe the redshift dependence of the AP effect, 
and we arrive at tight cosmological constraints competitive with mainstream probes of SNIa, BAO, CMB.

@@

When the correct cosmology is adopted the cosmic scale is correctly estimated,
the main factor leads to redshift evolution of galaxy distribution is the gravitational growth of structure.
With decreasing redshift, we clearly see the clusters and filaments form and grow.
On the other hand, when the wrong cosmology is adopted, there is an artificial scaling of distances in the constructed map.
The separations among galaxies are over-estimated by 
$25.6\%,47.3\%,62.2\%,71.7\%$ at redshifts of $0.5,1.0,1.5,2.5$,
leading to a clear redshift evolution of sizes of structures.

@@

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig3.eps}
    \includegraphics[width=18cm]{fig3_b.eps}
   }
   \caption{\label{fig_diffz}
  The angular 2pCF ($\omega_{r_\perp}$, left panels), the shape of the angular 2pCF ($\hat{\omega}_{r_\perp}$, middle panels),
  and the evolution with respect to the $z=0$ value ($\hat{\omega}_{r_{\perp}}(z=z_i) - \hat{\omega}_{r_{\perp}}(z=0)$, right panels). 
  Upper panels show that in the correct cosmology the redshift evolution of the shape of 2pCF is rather small regardless of the large redshfit evolution of the amplitude, 
  whereas the lower panels show that, in a wrong cosmology $\Omega_m = 0.05,w=-1.5$, the redshift evolution of $\hat{\omega}_{r_\perp}$ is very significant.
   }
\end{figure*}

The upper-left panel of Figure \ref{fig_diffz} displays the angular 2pCF measured from HR4 mock galaxies. % (multiplied by scale $r$) in the five snapshot.
We multiply $\omega$ by the separation $r_\perp$ to have similar size of statistical uncertainty on all scales.
For convenience we will use the denotation
\begin{equation}
 \hat \omega_{r_{\perp}} \equiv r_{\perp} \omega.
\end{equation}

Among different redshifts there is large variation in the amplitude of $\omega(r_\perp)$.
The amplitude is proportional to the clustering strength, which is affected by the gravitational growth of structures and the galaxy bias.
The amplitude is highest at $z=0$ where the structures experienced the most growth.
$\omega(r_\perp)$ increases at $z>1$ with increasing redshift 
as a reason of increasing galaxy bias.
%leading to stronger clustering strength and higher amplitude of 2pCF.
%On the contrary, at higher redshift we find $\omega(z=2)>\omega(z=1.5)>\omega(z=1)$,
%The trend is different from $z=1,1.5,2$,
%simply because by applying a uniform mass cut we are selecting more biased galaxies at higher redshift.

Although there is large variation in amplitude between the different redshifts bins, 
%The shape of $r_\perp\omega$ also evolves with redshift.
the shape of $\omega_{r_{\perp}}$  remains similar at all redshifts;
in general it peaks at $r\sim9 h^{-1}\rm Mpc$ and monotonically drops or increases at larger or smaller scales.
The only exception is the small enhancement at $r\lesssim 2 h^{-1}\rm Mpc$,
which is caused by the non-linear growth of structures 
and is much more significant at lower redshift.

@@

In the upper-middle panel, the overlapping of $\hat\omega_{r_{\perp}}$
clearly show the minimal redshift evolution of the shape.
In the upper-right panel, 
we further show the residual evolution at high redshifts 
with respect to $z=0$.
There is only 1-4\% enhancement 
at $r<10h^{-1}\rm Mpc$,
and $<1.5\%$ suppression at $r>25 h^{-1}\rm Mpc$.
The trend is monotonic with redshift.

@@

\subsection{Galaxy angular 2pCF: cosmological effect }

The angular 2pCF of galaxy positions constructed in a wrongly assumed cosmology $\Omega_m=0.05,\ w=-1.5$
is displayed in the lower panels of Figure \ref{fig_diffz}.
The mis-scaling uniformly shifts the clustering pattern on all scales, 
leading to a biased $\hat \omega_{r_\perp}$, which is 
related with the true $\hat \omega$ as
\begin{equation}
 \hat \omega_{r_\perp,\rm wrong}(r) = \hat \omega_{r_\perp, \rm correct}(\alpha_{\perp} r),
\end{equation}
a simple consequence of the fact that the clustering pattern at scale $r$ is rescaled to $\alpha_{\perp} r$.

The redshift evolution of $\alpha_{\perp}$ leads to redshift evolution of 
$\hat \omega_{r_\perp, \rm correct}$,
%Lower panels of Figure \ref{fig_diffz} shows the $r_\perp\omega$ in five 
%snapshots, in case that the cosmology $\Omega_m=0.05,\ w=-1.5$ 
%is adopted to construct the galaxy distribution (the right panels of Figure \ref{fig_scatter}).
which is displayed in the lower-middle panel.
The upscaling of comoving distance leads to a stretched shape.
When increasing redshift,
the peak location is shifted from $\sim9h^{-1} \rm Mpc$ to $\sim15 h^{-1}\rm Mpc$ at $z=2$,
as a result of the fact that the angular separation is upscaled by $71.7\%$.
Correspondingly, the right panel shows that, 
compared with $\hat \omega_{r_\perp}$ at $z=0$ there is a 20-40\% change at higher redshifts.

@@

We vary the mass cut and check the effect of galaxy bias on $\hat \omega_{r_\perp}$.
The upper panels of Figure \ref{fig_sys} displays the 2pCF measured from galaxies within a $3175\times3175\times105$ $(h^{-1}\rm Mpc)^3$ volume, taken from the $z=0$ snapshot.
Four values of minimal mass limits, $3\times 10^{11} M_{\odot}$, $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$,
are imposed to create samples with different galaxy bias.
The measured angular 2pCF are displayed.

Different mass cuts result in large variation in the amplitude of the 2pCF (left panel);
but the shape of the \scr{normalised} 2pCF remains less affected (middle panel).
Compared with the sample of $M>3\times 10^{11} M_{\odot}$,
samples with mass limits of $1\times 10^{12} M_{\odot}$, $4\times 10^{12} M_{\odot}$ and $1\times 10^{13} M_{\odot}$
have amplitude of 2pCFs enhanced by $10\%,\ 50\%$ and $100\%$, 
while the change in the shape is only $0.5\%,\ 2\%$ and $4\%$, respectively.
The effect of galaxy bias becomes much less significant by utilizing the shape of the 2pCF rather than the amplitude.

@@

The effect of RSD is much milder in this work. 
The angular positions of galaxies are not shifted by RSD at all;
the only effect of RSD enter in the procedure of 
splitting galaxies into redshift shells.
The galaxies observed in a survey are split into shells of subsamples, 
with different redshift ranges. So that we can get the 2pCF measurement at different redshfits.
RSD distorts the galaxy redshift and as a result some galaxies 
(especially those close to the boundaries of shells) are assigned to the wrong redshift shells.

We shift the radial coordinates of galaxies according to the relation 
\begin{equation}\label{eq:zvpeu}
\Delta z = (1+z) \frac{v_{{\rm Z}}}{c},
\end{equation}
where $v_{{\rm Z}}$ (Z is the third coordinate of the galaxy in the box, treated as the radial direction in this analysis) 
is the galaxy peculiar velocity in the LOS direction.
This lead to {\it misclassification} of some galaxies when we split the box into slices.
%Then we split the whole box into 30 slices with thickness $105h^{-1}\rm Mpc$, 
%and measure the angular 2pCF.

The lower panel of Figure \ref{fig_sys} shows how it affects on the 2pCF.
Comparing this plot with Figure \ref{fig_diffz} one can clearly see the effect of RSD.
The amplitude of the measured 2pCF is enhanced by $\sim 10\%$ when considering the RSD effect 
and the slop of $\hat\omega_{r_\perp}$ is suppressed.
But the redshift evolution of $\hat\omega_{r_\perp}$ is still small, remain $\lesssim4\%$ at $0.5\leq z \leq 2$.
So RSD should not very affect the application of our method as long as its redshift evolution is small;
the small effect of RSD can be precisely modeled by simulations and corrected.
%If looking at the redshift evolution,
%the curves $\hat{r_\perp\omega} - \hat{r_\perp\omega}(z=0)$ displayed in the right panel,
%they changed a lot compared with the case of no RSD effect (the right panel of Figure \ref{fig_diffz}),
%but do not become much larger. %But RSD effect does not introduce large redshfit evolution, although the shape of  is

Similar to RSD, redshift errors of galaxies
also results in fuzzy boundaries of shells and misclassification of galaxies.
This effect should be properly quantified when dealing with real observational data.

@@

Adopting the correct cosmological model will resulting in minimal evolution of the normalised 2pCF.
Thus we proceed to construct a quantitative likelihood estimator.
The procedure here is very similar to \cite{Li2014,Li2015,Li2016}.
We compare the high redshift $\hat\omega_{r_\perp}$ and the lowest redshift measurement,
\begin{equation}
 \delta \hat{\omega_{r_\perp}}(z_i,z_1) \equiv \hat{\omega_{r_\perp}}(z_i) - \hat{\omega_{r_\perp}}(z_1),
\end{equation}
and design the likelihood function to quantify
the redshift evolution of $\hat\omega_{r_\perp}$:
\begin{equation}\label{eq:chisq1}
\chi^2\equiv \sum_{i=2}^{n_{z}} \sum_{j_1=1}^{n_{r}} \sum_{j_2=1}^{n_{r}} {\bf p}(z_i,r_{j_1}) ({\bf Cov}_{i}^{-1})_{j_1,j_2}  {\bf p}(z_i,r_{j_2}),
\end{equation}
where $n_z$ is the number of redshifts which is 5 in this analysis, 
$n_r$ is number of binning in $\hat{\omega}_{r_\perp}(r_\perp)$,
which is 35 since we have $r_\perp$ bins 
with width $1h^{-1}\rm Mpc$ in a range of $5h^{-1}{\rm Mpc}\leq r_\perp\leq 40h^{-1}{\rm Mpc}$.
${\bf p}(z_i,r_{j})$ is the redshift evolution of the angular correlation function shape with systematic effects subtracted
\begin{eqnarray}\label{eq:bfp}
 {\bf p}(z_i,r_{j}) \equiv&\ \delta \hat{\omega_\perp}(z_i,z_1,r_j) - \delta \hat{\omega}_{r_\perp,\rm sys}(z_i,z_1,r_j)
\end{eqnarray}
${\bf Cov}_i$ is the covariance matrix estimated from the ${\bf p}(z_i,r_{j})$ measured from 120 subsamples.
%It is difficult to conduct the 2pCF of the whole snapshot box with hundreds of millions of galaxies.
%We split the box into 120 subsamples with $X,Y,Z$ sizes of $1575h^{-1}{\rm Mpc},1575h^{-1}{\rm Mpc}, 105h^{-1}{\rm Mpc}$
%and measure the angular 2pCF in the $X-Y$ plan in these subsamples. 
%The average of the 2pCFs are taken as the 2pCF of the whole box.
%The covariance matrix is also estimated from the redshift evolution of these subsamples.
As mentioned previously, for a robust estimation of the covariance matrix, 
we always compare slices at different locations to 
include the cosmic variance.

@@

Figure \ref{fig_cosmo} displays the 2pCF and the likelihood values when adopting 
the four incorrect cosmologies used in Figure \ref{fig_xyquan}.
For the cosmologies $\Omega_m=0.4,w=-1$ and $\Omega_m=0.26,w=-0.5$,
the compression shifts the clustering patterns into smaller scales
at higher redshift and make the 2pCFs steeper.
%and also becomes more significant at higher redshift.
%We get 2pCFs with steeper slope and the higher redshift the steeper.
For cosmologies $\Omega_m=0.15,w=-1$ and $\Omega_m=0.26,w=-1.5$,
the stretch of structure leads to shallower slope of $\hat{\omega}_{r_\perp}$ at higher redshift.
%the shallower.
In all cosmologies there is significant detection of redshift evolution.
We compute the $\chi^2$ values according to Equation \ref{eq:chisq1}, 
and find that these cosmologies are disfavored 
at $\gtrsim30\sigma$ CL.

\begin{figure*}
   \centering{
%   \includegraphics[width=18cm]{fig9_0.eps}
%   \includegraphics[width=18cm]{fig9_1.eps}
%   \includegraphics[height=8cm]{Tpcf--plot--Normed.eps}
%    \includegraphics[height=8cm]{smu.eps}
    \includegraphics[width=18cm]{fig5.eps}
   }
   \caption{\label{fig_cosmo}
    $\hat{\omega}_{r_\perp}$ (upper panels) and $\hat{\omega}_{r_\perp}(z=z_i) - \hat{\omega}_{r_\perp}(z=0)$ (lower panels) at several redshifts,
    for four wrong cosmologies.
    The redshift evolution of the angular 2pCF shape in these wrong cosmologies are detected at high CL.
   }
\end{figure*}
